---
title: 'Comparison of Global vs. Local Reference Databases and Taxonomic Cutoff Scores for Metabarcoding Taxonomic Assignment: Case Study on MiFish 12S Specific Reference Databases and California Current Large Marine Ecosystem Fishes'
author: "Zack Gold"
date: "12/4/2020"
output: html_document
---

```{r global-options, include=FALSE}
knitr::opts_chunk$set(warning=FALSE, message=FALSE)
```

```{r, echo=FALSE}
#Load Libraries
library(dplyr)
library(tidyverse)
library(knitr)
library(here)
```

```{r}
#Load Data

## California Fish Species
ca_fish_list <- read.table(here("Data","CA_fish_species","CA_fish_list_20210216.csv"), header = 1, sep = ",", stringsAsFactors = F)

## Species Monitored by PISCO, KFM, and RCCA
monitored <- read.table(here("Data","CA_fish_species","Monitored_Species_list_2020216.txt"), header = 1, sep = "\t", stringsAsFactors = F)

## Final Accessions
final_accessions <- read.table(here("Data","Supplemental Tables","final_accessions_20210521.txt"), header = 1, sep = "\t", stringsAsFactors = F)

## Local DB
FishCARD_db <- read.table(here("Data","local_database","Reference_db","trimmed","fishcard_12S_all_taxonomy_20201204.csv"), header = 1, sep = ",", stringsAsFactors = F)

FishCARD_db %>%
  filter(., str_detect(Final_name, "FISH")) %>% dplyr::select(Final_name) -> to_check_accessions 
setdiff(to_check_accessions$Final_name, final_accessions$Accession)
setdiff( final_accessions$Accession,to_check_accessions$Final_name)

FishCARD_db2 <- read.table(here("Data","local_database","Reference_db","trimmed","fishcard_12S_all_taxonomy.txt"), header = FALSE, sep = "\t", stringsAsFactors = F)

FishCARD_db2 %>%
  filter(., str_detect(V1, "FISH")) %>% dplyr::select(V1) -> to_check_accessions2
setdiff(to_check_accessions2$V1, final_accessions$Accession)
setdiff( final_accessions$Accession,to_check_accessions2$V1)

## Global DB
Global_db <- read.table(here("Data","global_database","Reference_db","trimmed","global_database_taxonomy_20201204.csv"), header = 1, sep = ",", stringsAsFactors = F)

Global_db %>%
  filter(., str_detect(Final_name, "FISH")) %>% dplyr::select(Final_name) -> to_check_accessions 
setdiff(to_check_accessions$Final_name, final_accessions$Accession)
setdiff( final_accessions$Accession,to_check_accessions$Final_name)

Global_db2 <- read.table(here("Data","global_database","Reference_db","trimmed","c19_fishcard_taxonomy.txt"), header = FALSE, sep = "\t", stringsAsFactors = F)

Global_db2 %>%
  filter(., str_detect(V1, "FISH")) %>% dplyr::select(V1) -> to_check_accessions2
setdiff(to_check_accessions2$V1, final_accessions$Accession)
setdiff( final_accessions$Accession,to_check_accessions2$V1)

```

# Barcoding Effort Statistics

#### List of All California Fish Species
```{r}
ca_fish_list$Walker_Miller_Lea_Combo_20210216 %>% unique() %>% length()
#1144 species in CA
```

#### CA Fish Species Barcoded by our Efforts
```{r}
FishCARD_db %>% 
  filter(., str_detect(Final_name,"FISHCARD_")) %>% 
  filter(., !str_detect(Species," sp[.]")) %>% 
dplyr::select(Species) %>% unique() -> fishcard_only_species

fishcard_only_species %>% dim()
#607 total species barcoded
```

#### Average Numbder of Barcodes per Species
```{r}
FishCARD_db %>% 
  filter(., str_detect(Final_name,"FISHCARD_")) %>% 
dplyr::select(Species) %>% 
group_by(Species) %>% 
  dplyr::count() %>% ungroup() %>% 
  dplyr::summarise(mean(n), max(n), min(n))
```

#### Average Numbder of Barcodes per Species
```{r}
FishCARD_db %>% 
  filter(., str_detect(Final_name,"FISHCARD_")) %>% 
  filter(., str_detect(Final_name,"elas")) %>% 
dplyr::select(Species) %>% 
group_by(Species) %>% 
  dplyr::count() %>% ungroup() %>% 
  dplyr::summarise(mean(n), max(n), min(n))
```

#### Average Numbder of Barcodes per Species
```{r}
FishCARD_db %>% 
  filter(., str_detect(Final_name,"FISHCARD_")) %>% 
  filter(., str_detect(Final_name,"miu")) %>%
dplyr::select(Species) %>% 
group_by(Species) %>% 
  dplyr::count() %>% ungroup() %>% 
  dplyr::summarise(mean(n), max(n), min(n))
```

### Species in Global
```{r}

Global_db %>% 
  dplyr::select(Species) %>%  unique() %>%  dim()

Global_db %>% dim()

```

#### Teleost in Local DB
```{r}
FishCARD_db %>% 
  filter(., str_detect(Final_name,"FISHCARD_")) %>% 
  filter(., Class=="Actinopteri") %>% 
    filter(., !str_detect(Species," sp[.]")) %>%
  dplyr::select(Species) %>% unique() %>% dim()
#552 Bony Fish species
```


#### Sharks and Rays in Local DB
```{r}
FishCARD_db %>% 
  filter(., str_detect(Final_name,"FISHCARD_")) %>% 
  filter(., Class=="Chondrichthyes") %>% 
    filter(., !str_detect(Species," sp[.]")) %>% 
  dplyr::select(Species) %>% unique() %>% dim()
#49 shark Fish species
```

#### Boneless Fish Species
```{r}
FishCARD_db %>% 
  filter(., str_detect(Final_name,"FISHCARD_")) %>% 
  filter(., Class %in% c("Myxini","Hyperoartia")) %>% 
  dplyr::select(Species) %>% unique() %>% dim()

#4 boneless Fish species
```

#### Check CA Fish Species List
```{r}
FishCARD_db %>% 
  filter(., str_detect(Final_name,"FISHCARD_")) %>% 
  dplyr::select(Species) -> FishCARD_df
setdiff(FishCARD_df$Species,ca_fish_list$Walker_Miller_Lea_Combo_20210216) -> barcodes_with_genus_level_ids
barcodes_with_genus_level_ids
## 2 taxa barcoded could only be identified to Genus level, will remove for calculating statistics
```

#### List of All CA Fish Species that Now Have MiFish *12S* References
```{r}

FishCARD_db %>% 
  dplyr::select(Species) %>% 
  filter(., !Species %in% barcodes_with_genus_level_ids) -> FishCARD_df_total

FishCARD_df_total$Species %>% unique() -> local_db_species_list
local_db_species_list %>%  length()

write.csv(local_db_species_list, file =here("Data","CA_fish_species","CA_fish_covered.csv"))

```

#### CA Fish Species that are still missing MiFish Reference Barcodes
```{r}

setdiff(ca_fish_list$Walker_Miller_Lea_Combo_20210216,local_db_species_list) -> Ca_fish_missing_mifish_barcodes

setdiff(local_db_species_list,ca_fish_list$Walker_Miller_Lea_Combo_20210216)

local_db_species_list %>% length()

Ca_fish_missing_mifish_barcodes %>%  length()
ca_fish_list$Walker_Miller_Lea_Combo_20210216 %>% length()
local_db_species_list %>%  length()

write.csv(Ca_fish_missing_mifish_barcodes,file=here("Data","CA_fish_species","Ca_fish_missing_mifish_barcodes.csv"))
```

#### Percent of CA Fish species Now Barcoded

```{r}
length(local_db_species_list)/length(ca_fish_list$Walker_Miller_Lea_Combo_20210216)*100

```


